---
title: "Oxygen Saturation Regression Model"
author: "Kiera Matiska"
date: "1/20/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(AICcmodavg)
library(equatiomatic)
library(kableExtra)
```

## Read in data

```{r}
seawater_chem <- read_csv(here("data", "calcofi_seawater_samples.csv")) # read in data
```

## Create two multiple linear regression models

- Model 1: Oxygen saturation as a function of water temperature, salinity, and phosphate concentration
- Model 2: Oxygen saturation as a function of water temp, salinity, phosphate concentration, and depth.

```{r}
f1 <- o2sat ~ t_deg_c + salinity + po4u_m # first model
seawater_mdl1 <- lm(f1, data = seawater_chem) # multiple linear regression of model

f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m # second model
seawater_mdl2 <- lm(f2, data = seawater_chem) # multiple linear regression of model
```

### Compare AIC values for the two models

```{r}
AIC(seawater_mdl1, seawater_mdl2) # 618.3868, 615.7016; calculate AIC values

AICc(seawater_mdl1) # 619.0251; calculate AICc value for model 1
AICc(seawater_mdl2) # 616.6048; calculate AICc value for model 2
aictab(list(seawater_mdl1, seawater_mdl2)) %>% # identify difference between the AIC values
  kable() %>% # create a finalized table with the values
  kable_styling(position = "center")
```

Based on the AIC scores, the second model is the better out of the two. The difference between the two models is `r 619.0251 - 616.6048`.

### K-fold Cross Validation

```{r}
n_folds <- 10 # define 10 folds
fold <- rep(1:n_folds, length.out = nrow(seawater_chem)) 
table(fold) # create a table with the fold values

set.seed(42) # set seed to continue to keep values the same

seawater_fold <- seawater_chem %>% # create a subset
  mutate(group = sample(fold, size = n(), replace = FALSE)) # add a new group column

# first fold
seawater_test_df <- seawater_fold %>% 
  filter(group == 1)
seawater_train_df <- seawater_fold %>% 
  filter(group != 1)
```

Calculate root mean square

```{r}
# root mean square = (x-y)^2, take the mean, and then square root the mean number
calc_rmse <- function(x, y) {
  rmse_result <- (x - y)^2 %>% mean() %>% sqrt()
  return(rmse_result)
}
```

Create training data sets using the two equations above

```{r}
training_mdl1 <- lm(f1, data = seawater_train_df) # create training model for model 1
training_mdl2 <- lm(f2, data = seawater_train_df) # create training model for model 2
```

Use trained models to predict on test data

```{r}
# create a predicted data table with predictions for model 1 and 2
seawater_predict <- seawater_test_df %>% 
  mutate(model1 = predict(training_mdl1, seawater_test_df),
         model2 = predict(training_mdl2, seawater_test_df))

seawater_rmse <- seawater_predict %>% 
  summarize(seawater_rmse_mdl1 = calc_rmse(model1, o2sat),
            seawater_rmse_mdl2 = calc_rmse(model2, o2sat))

seawater_rmse %>%  
  kable() %>% 
  kable_styling(position = "center")
```

Let's calculate over all folds and take the average

```{r}
seawater_rmse_df <- data.frame()

for(i in 1:n_folds) {
  kfold_test_df <- seawater_fold %>% 
    filter(group == i)
  kfold_train_df <- seawater_fold %>% 
    filter(group == i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .))
  kfold_rmse <- kfold_pred_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, o2sat),
              rmse_mdl2 = calc_rmse(mdl2, o2sat))
  seawater_rmse_df <- bind_rows(seawater_rmse_df, kfold_rmse)
}

seawater_rmse_df %>% 
  summarize(mean_rmse_mdl1 = mean(rmse_mdl1),
            mean_rmse_mdl2 = mean(rmse_mdl2))
```

Model 2 has both the lowest RMSE and AIC values, therefore it is the better model out of the two.

## Final Model Choice After Cross-Validation and AIC calculations

```{r}
final_seawater_mdl <- lm(f2, data = seawater_chem)
summary(final_seawater_mdl)
```

Final Equation: `r extract_eq(final_seawater_mdl, wrap = TRUE, use_coefs = TRUE)`

Your organized code, with clear subsections and any useful descriptive text / annotation (e.g. if you wanted to highlight this as a code example for a prospective employer).  **Include an overview of the data and your exploration, and a citation for the data.  Suppress all warnings and messages, but show all attached packages.**
Your comparison of the two (or three?) models using an appropriate application of AIC, with the resulting AIC scores and an interpretation of how to choose the best model with this information.  **Communicate results in a nicely formatted manner (in-line code chunks, kable or other formatted table, etc).**
Your comparison of the two (or three?) models using a 10-fold cross validation, with the resulting average RMSE value across all folds, and an interpretation of how to choose the better model with this information.  **Communicate results in a nicely formatted manner (in-line code chunks, kable or other formatted table, etc).**
The final parameterized model (including coefficient values) expressed as a formatted equation (using equatiomatic or LaTeX).  **If your two model selection methods identify different models, select the result of one method and explain why you prefer that method.**

