---
title: "Oxygen Saturation Regression Model"
author: "Kiera Matiska"
date: "1/20/2022"
output: html_document
---
# Overview:

Ocean chemistry is important for monitoring ocean health. This data analysis looks to find the model that gives the most accurate O_2 saturation level based on different chemistry factors in the ocean. The data used for this analysis is from seawater off the coast of California, and levels have been recorded since 1949. The variables measured for this data include oxygen saturation levels, temperature in degrees Celsius, salinity, depth in meters, chlorophyll-a concentrations measured in micro-grams per liter, phosphate concentration in micro-moles per liter, and nitrite concentrations in micro-moles per liter. For this analysis, two multiple linear regression models will be generated to calculate Oxygen Saturation levels, and the models will be compared using AIC values and cross-validation.

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(AICcmodavg)
library(equatiomatic)
library(kableExtra)
```

## Read in data

```{r}
seawater_chem <- read_csv(here("data", "calcofi_seawater_samples.csv")) # read in data
```

## Create two multiple linear regression models

- Model 1: Oxygen saturation as a function of water temperature, salinity, and phosphate concentration
- Model 2: Oxygen saturation as a function of water temperature, salinity, phosphate concentration, and depth.

```{r}
f1 <- o2sat ~ t_deg_c + salinity + po4u_m # first model
seawater_mdl1 <- lm(f1, data = seawater_chem) # multiple linear regression of model

f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m # second model
seawater_mdl2 <- lm(f2, data = seawater_chem) # multiple linear regression of model
```

### Compare AIC values for the two models

```{r}
AIC(seawater_mdl1, seawater_mdl2) # 618.3868, 615.7016; calculate AIC values

AICc(seawater_mdl1) # 619.0251; calculate AICc value for model 1
AICc(seawater_mdl2) # 616.6048; calculate AICc value for model 2
```

**Table 1.** AIC values for the two different models and the difference between the two values
```{r}
aictab(list(seawater_mdl1, seawater_mdl2)) %>% # identify difference between the AIC values
  kable() %>% # create a finalized table with the values
  kable_styling(position = "center") # center the table
```

Based on the AIC scores, the second model is the better out of the two. The difference between the two models is `r 619.0251 - 616.6048`.

### K-fold Cross Validation

```{r}
n_folds <- 10 # define 10 folds
fold <- rep(1:n_folds, length.out = nrow(seawater_chem)) 
table(fold) # create a table with the fold values

set.seed(42) # set seed to continue to keep values the same

seawater_fold <- seawater_chem %>% # create a subset
  mutate(group = sample(fold, size = n(), replace = FALSE)) # add a new group column

# first fold
seawater_test_df <- seawater_fold %>% 
  filter(group == 1)
seawater_train_df <- seawater_fold %>% 
  filter(group != 1)
```

Function for Root-Mean-Square Values

```{r}
# root mean square = (x-y)^2, take the mean, and then square root the mean number
calc_rmse <- function(x, y) {
  rmse_result <- (x - y)^2 %>% mean() %>% sqrt()
  return(rmse_result)
}
```

Create training data sets using the two equations above

```{r}
training_mdl1 <- lm(f1, data = seawater_train_df) # create training model for model 1
training_mdl2 <- lm(f2, data = seawater_train_df) # create training model for model 2
```

Use trained models to predict on test data

```{r}
# create a predicted data table with predictions for model 1 and 2
seawater_predict <- seawater_test_df %>% 
  mutate(model1 = predict(training_mdl1, seawater_test_df),
         model2 = predict(training_mdl2, seawater_test_df))

# find RMSE values for the two models using prediction data
seawater_rmse <- seawater_predict %>% 
  summarize(seawater_rmse_mdl1 = calc_rmse(model1, o2sat),
            seawater_rmse_mdl2 = calc_rmse(model2, o2sat))
```

**Table 2.** Predicted RMSE values for models 1 and 2 based on the test data frame.
```{r}
seawater_rmse %>%  
  kable() %>% # create a table with this information
  kable_styling(position = "center") # center the table
```

Calculations over all folds and take the average

```{r}
seawater_rmse_df <- data.frame()

for(i in 1:n_folds) {
  kfold_test_df <- seawater_fold %>% 
    filter(group == i)
  kfold_train_df <- seawater_fold %>% 
    filter(group == i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .))
  kfold_rmse <- kfold_pred_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, o2sat),
              rmse_mdl2 = calc_rmse(mdl2, o2sat))
  seawater_rmse_df <- bind_rows(seawater_rmse_df, kfold_rmse)
}

seawater_rmse_df %>% 
  summarize(mean_rmse_mdl1 = mean(rmse_mdl1),
            mean_rmse_mdl2 = mean(rmse_mdl2))
```

**Table 3.** RMSE values for Models 1 and 2.
```{r}
kfold_rmse %>% 
  kable() %>% 
  kable_styling(position = "center")
```

Based on this cross validation, the root-mean-square of model 1 is `r round(kfold_rmse$rmse_mdl1[1], 2)`. Model 2's root-mean square value is `r round(kfold_rmse$rmse_mdl2[1], 2)`. Since model 2 has the smallest root-square value, it is the better model of the two.

Based on the AIC values and root-mean-square values of model 1 and model 2, model 2 is the best because it has both the smallest AIC value and root-mean-square value.

## Final Model Choice After Cross-Validation and AIC calculations

```{r}
final_seawater_mdl <- lm(f2, data = seawater_chem)
summary(final_seawater_mdl)
```

Since Model 2 is the better model, the final equation is:

`r extract_eq(final_seawater_mdl, wrap = TRUE, use_coefs = TRUE)`

#### Data Citation

CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.  Accessed 1/20/2022.
